Episodic memory: mental time travel or a quantum `memory wave' function?

### 2020-09-19 paper/main.tex
working on word embedding section

% topic models-- deeper themes, stress interpretability.  example: bat (animal, sports equipment, verb).  "As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.  The swing connected, stunning the creature before it scrambled about on the ground and flew away.  She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity."

%deep learning approaches

>>>>

\textit{Topic models}, such as \textit{Latent Dirichlet Allocation}~\citep[LDA;][]{BleiEtal03} represent an attempt to characterize the latent \textit{topics} or themes that documents comprise (Fig.~\ref{fig:embedding-models}B).  Whereas matrix factorization approaches represent semantic meanings using low-dimensional approximations of word co-occurances or free association frequencies, topic models leverage these properties of documents (or behaviors) to explain \textit{why} words co-occur.  LDA defines a topic as a distribution of weights over words in the vocabulary.  For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic would comprise a vector of 10,000 numbers (weights) between zero and one, inclusive, where the full set of weights summed to one.  As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five topics are displayed), a fruit-related topic (shown in red) might weight words like ``apple,'' ``banana,'' ``pear,'' and ``orange'' heavily, whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.  Given a set of $K$ topic vectors, topic models compute the blend of topics most probably reflected in a given document-- i.e., its \textit{topic proportions}.  These topic proportion estimates are made by examining how different topics weight each of the words in the document, and by assuming that documents tend to reflect few topics (rather than many topics).  Given only a word count matrix as input (just as in LSA), topic models automatically discover a $K$ by number-of-words \textit{topics matrix}, which describes how each topic (row) weights each word in the vocabulary (column), and a number-of-documents by $K$ \textit{topic proportions matrix}, which describes each document's topic proportions.  The rows of the topic proportions matrix may be treated as low-dimensional embeddings of the set of documents.

A key advantage that topic models have over matrix factorization models is the ability to explain potentially ambiguous word usage.  For example, consider the use of the word ``bat'' in the following passage:
\begin{quote}
  \textit{``...As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.  The swing connected, stunning the creature before it scrambled about on the ground and flew away.  She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity...''}
  \end{quote}
At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and a verb.  Because each of these uses of the word appear in the same document, models driven solely by word co-occurance statistics will blend these different meanings such that the embedding of ``bat'' is somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.  By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.  If other documents that mention the word ``bat'' describe vampires, swooping, or other mammals, but \textit{not} sports equipment or eye-related verbs, then a topic model could explain this observation by inferring that multiple topics place substantial weight on the word ``bat.''  Another strength of topic models is their interpretability.  Because the models directly output each document's topic proportions along with each topic's weights on all words in the vocabulary, the general themes that each topic reflects are often intuitive and the per-document topic proportions may be directly related to those themes.

% deep learning approaches
% bag of words vs. sequence models, context and grammar to convey meaning, etc.
